General constraints on magic distillation protocols from statistical mechanics

### 2021-04-29 magic_paper.tex
Added proof with one comment/question on a part I am not sure about. Pushed Davids text to notes.tex. Need to fix text following theorem / describing leaf plot

The Wigner distribution of $n$ copies of the state is obtained by the $n$-fold product of the single copy distribution, and the Lorenz curve $L_{\rho|\sigma}(x)$ for $\rho = \rho_S(\epsilon)^{\otimes n}$ and $\sigma = (\I/3)^{\otimes n}$ is readily computed for any $n$ and $\epsilon$, and the full details are provided in~\cref{app:lcsu_technical}.
We highlight that the non-trivial element that arises for quasi-distributions is that the product of negative components can generate a positive $n$-fold component, so the ordering of the components of $W_{\rho |\sigma}(x)$ depends on whether $-v(\epsilon)$ appears to an even or odd power in the expression of~\cref{eq:wigu}.

@@

The bounds also depend on the overlaps between the Hamiltonian eigenbases and the computational basis $\{\ket{k}\}$, where $k \in \mathbb{Z}_d$.
For a reference state $\tau$ with associated Hamiltonian $H$ denote the energy eigenbasis by $\{\ket{\phi_k}\}$ and define the \emph{reference partition function $\zeta$} as
\begin{equation}
	\zeta \coloneqq \sum_{k \in \mathbb{Z}_d} \alpha_k e^{-\beta E_k},
\end{equation}
where the overlap coefficients are
\begin{equation}
	\alpha_k \coloneqq \sum_{\mathbb{Z}_d} \braket{\phi_k}{-r}\braket{r}{\phi_k}.
\end{equation}
Coefficients $\alpha_k$ are not necessarily positive, which is the case when the Hamiltonian $H$ does not admit a stabilizer decomposition. 
However, it is simple to see that $\zeta$ is always positive. \nick{because it is proportional to $W_{\tau}(0,0) > 0$ for full-rank, stabilizer $\tau$ - this argument is a bit cyclic}
We can then naturally associate to the reference state $\tau$ a new form of free energy which is specific to the protocol, which we call the \emph{reference free energy $\Phi$} and define via
\begin{equation}
	\Phi \coloneqq -\beta^{-1} \log \zeta.
\end{equation}
For the output reference state $\sigma$ with associated Hamiltonian $H'$, we can denote the energy eigenbasis by $\{\ket{\phi'_k}\}$ and similarly define the reference partition function $\zeta'$ and reference free energy $\Phi'$.

@@

Then, the distillation rate $R = m/n$ of the magic protocol is bounded as:
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\Phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\Phi' - F')},
\end{equation}
where $F, \Phi$ are the free energy and reference free energy defined for state $\tau$, while $F', \Phi'$ are defined for state $\sigma$.
\end{theorem}
\begin{proof}
	\nick{Start}
\end{proof}

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

Let each qutrit have a Hamiltonian $H$ with stabilizer eigenstates and energies $E_0, E_1, E_2$, and define $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and $E_s$ the eigenvalue of the eigenstate that \nick{overlaps} the negative component of $\rho_S(\epsilon)$ in the Wigner representation. Let $T =(k\beta)^{-1}$ be any characteristic temperature for the physical system in the state $\tau^{\otimes n}= (e^{-\beta H}/\Z)^{\otimes n}$, with free energy $F$. Assume that for $n,m \gg 1$ the channel $\E$ generates negligible correlations on $\tau^{\otimes n}$, and so $\E(\tau^{\otimes n}) = \sigma^{\otimes m}$ for some state $\sigma$.

Define $\beta_\star = (k T_\star)^{-1}$ through the relation
\begin{equation}
	E_{\rm max} - E_s \eqqcolon kT_\star \ln 2,
\end{equation}
and define a threshold noise,
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 
	\begin{cases}
		3 - \dfrac{9}{4-2^{\beta/\beta_\star - 1}}, &\text{ for } \beta \leq \beta_\star \\
		0, &\text{ for } \beta > \beta_\star.
	\end{cases}
\end{equation}
Then, the distillation rate $R = m/n$ of the magic protocol is bounded as:
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\epsilon \leq \epsilon_\star, \epsilon' \leq \epsilon'_\star, \vspace{10pt}\\
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\epsilon \leq \epsilon_\star, \epsilon' > \epsilon'_\star, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\epsilon > \epsilon_\star, \epsilon' \leq \epsilon'_\star, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\epsilon > \epsilon_\star, \epsilon' > \epsilon'_\star,
	\end{cases}
\end{equation}
where $F'$ is the free energy of the state $\sigma = e^{-\beta H'}/\Z'$ and other primed quantities are defined in the same way as the corresponding unprimed quantities. 
\end{theorem}

We note, firstly, that the specific numerical factors in $\epsilon_\star$ are a result of our choice of magic state. 
Secondly, these bounds are derived based on analysis of only a part of the Lorenz curves and can be improved via a finer analysis. 
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

As a special case of the result, suppose that $H=H'$, implying that the protocol leaves the equilibrium state unchanged, hence it corresponds to a Gibbs-preserving map~\cite{faist_2015}. In the more general case, $\Delta F \coloneqq F' - F \ne 0$ and we have a free energy difference between the initial and final states, with respect to the relevant Hamiltonians. This is similar to the free energy differences that occur in fluctuation relations, where the final state of the system can be arbitrarily far from equilibrium, yet a free energy term for the final system still arises.

The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.45]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\begin{proof}
To establish the distillation bounds we consider the distillation protocol that gives
\begin{equation}
	\E( \rho_S(\epsilon)^{\otimes n}) = \rho_S(\epsilon')^{\otimes m},
\end{equation}
for the magic state. We then consider how the protocol transforms the reference equilibrium state as
\begin{equation}
	\E( \tau^{\otimes n}) = \sigma^{\otimes m}.
\end{equation}
Since $\tau$ and $\sigma$ are assumed to be full rank stabilizer states they have a strictly positive Wigner distribution, while, in contrast, the input and output magic states will generally have quasi-probability distributions for their Wigner functions. For any such protocol we therefore have that
\begin{equation}
	(W_{\rho_n} (x), W_{\tau_n}(x) ) \succ  (W_{\rho'_m} (x), W_{\sigma_m}(x) ),
\end{equation}
where for simplicity we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_m\coloneqq \rho_S(\epsilon')^{\otimes m}$ and $\tau_n\coloneqq\tau^{\otimes m}$ and $\sigma_m\coloneqq\sigma^{\otimes m}$. We then have that
\begin{equation}
	L_{\rho_n |\tau_n}(x) \ge L_{\rho'_m |\sigma_m}(x) \mbox{ for all } x.
\end{equation}
We must therefore compute the Lorenz curve data for $\rho_n$ relative to $\tau_n$, and compare with the Lorenz curve of the output state $\rho_m'$ relative to $\sigma_m$.

We define $W_{\rho | \tau}(\x) \coloneqq W_\rho(\x)/W_\tau(\x)$, which is always well-defined since $\tau$ is full-rank. Now we show in~\cref{app:wigner} that
\begin{equation}
W_{\rho_1\otimes \rho_2 | \tau_1 \otimes \tau_2} (\x_1 \oplus \x_2) = W_{\rho_1 | \tau_1}(\x_1)W_{\rho_2 | \tau_2}(\x_2),
\end{equation}
and also,
\begin{equation}
W_{\rho_1\otimes \rho_2} (\x_1 \oplus \x_2) = W_{\rho_1}(\x_1)W_{\rho_2}(\x_2),
\end{equation}
for any states $\rho_1, \rho_2$ and any full-rank stabilizer states $\tau_1, \tau_2$. \ddd{[Notation clashing here. Frame result outside as its own lemma -- we gotta check everything.]} \nick{Will frame in~\cref{app:wigner} and fix notation in the next commit}
Therefore we have that
\begin{align}
W_{\rho_n |\tau_n} (\x) &= \prod_{i=1}^n W_{\rho|\tau}(\x_i)\\
W_{\rho_n } (\x) &= \prod_{i=1}^n W_\rho(\x_i)
\end{align}
where $\x = \oplus_{i=1}^n \x_i$ is the phase space point for the full system in terms of those of the individual subsystems.

The points defining the Lorenz curve $L_{\rho_n |\tau_n}(x)$ are obtained from first sorting the components of $W_{\rho_n |\tau_n}(\x)$ in non-increasing order and then computing the partial sums of $W_{\rho_n |\tau_n}(\pi(\x))$ where $\pi$ is the permutation that realises the sorting. Therefore, we first look at the values of $W_{\rho|\tau}(\x_i)$ for the single-copy case.

The equilibrium state at inverse temperature $\beta$ on a single qutrit is given by $\tau = e^{-\beta H} / \Z$. Moreover we have that $\tau$ is a full-rank stabilizer state, where $\beta \geq 0$ and $H = E_0 \ketbra{\varphi_0} + E_1 \ketbra{\varphi_1} + E_2 \ketbra{\varphi_2}$ is an eigendecomposition of $H$.
The state $\tau$ can now be written as 
\begin{equation}
	\tau = \frac{e^{-\beta E_0}}{\Z} \ketbra{\varphi_0} + \frac{e^{-\beta E_1}}{\Z} \ketbra{\varphi_1} + \frac{e^{-\beta E_2}}{\Z} \ketbra{\varphi_2},
\end{equation}
where the eigenstates $\{\ket{\varphi_k}\<\varphi_k|\}$ are pure, orthonormal stabiliser states, which can be represented in terms of generalized Paulis. To make our analysis simpler, we perform a change of basis that does not affect the Wigner negativity of the problem. We let $C$ be the unitary transforming each $|\varphi_k\>\<\varphi_k|$ to $|k\>\<k|$. Since the Clifford group is the normalizer of the Heisenberg-Weyl group, $C$ is a Clifford unitary. Therefore, $C$ maps $\tau$ to another stabilizer state that is diagonal in the computational basis, and we can assume without loss of generality that $\tau$ is diagonal in $\{|0\>,|1\>, |2\>\}$. However, this choice means that the location of the negative Wigner component $-v(\epsilon)$ of the Strange state will be permuted on the discrete phase space. We denote by $E_s$ the eigenvalue of $H$ where the associated eigenvector has Wigner distribution overlapping the negative component of the magic state $C\rho_S(\epsilon)C^\dagger$.  This is unique, since the eigenstates form an orthonormal basis.

The Wigner distribution of state $\tau$ is then given by
\begin{align}
	\W[\bmx]{\tau} &= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z}W_{\ketbra{k}}(x, p) \nonumber\\
	&= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z} \delta_{x, k} = \frac{e^{-\beta E_x}}{3\Z},
\end{align}
where $x$ labels one of the three vertical lines in the phase space.
The rescaled Wigner distribution $W_{\rho|\tau}(\x)$ is then easily computed. It has $9$ components, but several of these come with multiplicities. In total, there are four distinct values on the phase space, as illustrated in~\cref{fig:pd_split}.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{figs/pd_split_thermal.pdf}
    \caption{\textbf{Qutrit phase space regions for $W_{\rho | \tau}(\x)$.}
    Here, the negative component of the magic state overlaps the Wigner distribution of $|0\>$. The rescaled distribution attains a single value in each of the four regions, proportional to the value depicted in the region, see~\cref{eq:bmw_rescaled}.
    }
    \label{fig:pd_split}
\end{figure}

We now denote by $\w(\rho), \w(\rho|\tau)$ the unique values occurring in $W_\rho(\x), W_{\rho|\tau}(\x)$ respectively and $\bmm$ the vector of associated multiplicities of each value in $W_\rho(\x)$. The component values and multiplicities of the relevant distributions in the four distinct regions are given by
\begin{align}
	\w(\rho) &\coloneqq (-v, u, u, u), \\
		\bmm &\coloneqq (1,2,3,3), \\
	\bmw(\tau) &\coloneqq \frac{1}{3\Z} \left( e^{-\beta E_0}, e^{-\beta E_0}, e^{-\beta E_1}, e^{-\beta E_2} \right), \\
	\bmw(\rho_{\rm{S}} | \tau) &\coloneqq 3\Z \left( -v e^{\beta E_0}, u e^{\beta E_0}, u e^{\beta E_1}, u e^{\beta E_2} \right). \label{eq:bmw_rescaled}
\end{align}

Using this notation, the values and multiplicities of the $n$--copy distribution $\bmw(\rho_n |\tau_n)$ are computed in~\cref{lem:ncopycomponents} in~\cref{app:cmpairs}. The values are given by 
\begin{align}\label{eq:ncopy_w_rescaled}
	[\w(\rho_n | \tau_n)]_{ijk} &= (3\Z)^{n} (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )},
\end{align}
where the indices $i,j,k$ are non-negative integers that obey the constraint $\alpha \coloneqq i+j+k \leq n$.
The multiplicity of this above value is $m_{ijk}$ with
\begin{equation}
	m_{ijk} = \frac{n!}{i!j!k!(n-\alpha)!} 2^i 3^j 3^k.
\end{equation}
The associated components of $\w(\rho_n)$ are given by
\begin{align}
	[\w(\rho_n)]_{ijk} &= (-v)^{n-\alpha} u^{\alpha}, \label{eq:ncopy_wrho}\\
	[\w(\tau)]_{ijk} &= (3\Z)^{-n} e^{-\beta (n-\alpha)E_s} e^{-\beta ( i E_0 + j E_1 + k E_2 )}. \label{eq:ncopy_wsigma}
\end{align}

In order to construct the $n$--copy Lorenz curve $L_{\rho_n|\tau_n}(x)$ we need to order the components of the distribution, $w(\rho_{\rm{S}} | \tau)_{ijk}$ in decreasing order, and identify the sequence of indices that give us $W_{\rho_n}(\pi(\x))$.

Generally this is complex, but in order to obtain distillation bounds it is sufficient to determine the location of the first elbow $(x_0, L_0)$ of $L_{\rho_n|\tau_n}(\x)$. To do so, we compute the largest component 
\begin{equation}
	w_{\rm max} \coloneqq \max_{i,j,k} [\w(\rho_n | \tau_n)]_{ijk},
\end{equation}
and determine the indices at which this occurs.
Putting in the values we obtain
\begin{align}
	&(3\Z)^{-n}w_{\rm max} = \nonumber\\
	&\max\limits_{i,j,k}\Big\{ (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )} \Big\}, \label{eq:max_slope}
\end{align}
where $0 \leq i,j,k \leq n$ and $\alpha \coloneqq i+j+k \leq n$.
Now for $0 \leq \epsilon \leq 3/7$, we have $v \geq u$. Since we assume that $n$ is even, we need the sum $\alpha = i+j+k$ to be even too, so that the expression is positive.

Given an even value for $\alpha$, the term $v^{n-\alpha} u^{\alpha} e^{-\beta (n-\alpha)E_s}$ is fixed, so the expression is maximised by setting the coefficient of the highest energy $E_{\rm{max}}$ equal to $\alpha$.
Hence, we have
\begin{align}
	&w_{\rm max} = \nonumber\\
	&(3\Z)^{n} v^n e^{n\beta E_s}\max\limits_{\substack{\alpha = 0,2, \\ \dots,n-2,n}}{\Big\{ \left( \frac{u}{v} e^{\beta (E_{\rm{max}} - E_s)} \right)^{\alpha} \Big\}}.
\end{align}
If the expression $\frac{u(\epsilon)}{v(\epsilon)} e^{\beta (E_{\rm{max}} - E_s)}$ is less than $1$ then the maximum occurs at $\alpha=0$, otherwise it occurs at $\alpha = n$. For a fixed state $\tau$, this transition is determined by the value of the depolarising noise parameter $\epsilon$ of the noisy magic state. The transition occurs at $\epsilon = \epsilon_\star$ where
\begin{equation}\label{eq:noise_transition}
	\frac{u(\epsilon_\star)}{v(\epsilon_\star)} e^{\beta (E_{\rm{max}} - E_s)} = \frac{3-\epsilon_\star}{6-8\epsilon_\star} e^{\beta (E_{\rm{max}} - E_s)} = 1.
\end{equation}
If $E_{\rm{max}} = E_s$, namely if the state negativity lies in the same phase space region as the highest energy, this threshold is constant in temperature and given by $\epsilon_{\star} = 3/7$. However, the condition that $\epsilon_\star \ge 0$ also implies a constraint on the effective temperature of the stabilizer state. Specifically, there is a threshold temperature value $\beta_\star$ given by
\begin{equation}
	\beta_{\star} \coloneqq \frac{1}{E_{\rm{max}} - E_s} \ln2,
\end{equation}
such that for the regime $0 \leq \beta \leq \beta_\star$ a transition noise level $\epsilon_\star$ exists, and for $\beta > \beta_\star$ no such transition exists, so we choose $\epsilon_\star = 0$. 
Therefore, the transition value for the noise is given by
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 
	\begin{cases}
		3 - \dfrac{9}{4-2^{\beta/\beta_\star - 1}}, &\text{ for } \beta \leq \beta_\star \\
		0, &\text{ for } \beta > \beta_\star.
	\end{cases}
\end{equation}
The quantity $w(\rho_{\rm{S}} | \sigma)_{\rm{max}}$ is now given by
\begin{equation*}
w_{\rm max} =
	\begin{cases}
		(3\Z)^{n} v^n e^{n\beta E_s}, &\mbox{if }\epsilon \leq \epsilon_{\star},\ \hspace{3pt}\rm{(C1)}\\
		(3\Z)^{n} u^n e^{n\beta E_{\rm{max}}}, &\mbox{if }\epsilon > \epsilon_{\star}.\ \hspace{5pt}\rm{(C2)} 
	\end{cases}
\end{equation*}
Case $\rm{(C1)}$ corresponds to $(i,j,k) = (0,0,0)$, so the multiplicity is $m_{000} = 1$, while
Case $\rm{(C2)}$ corresponds to
\begin{equation}
	(i,j,k) = 
	\begin{cases}
	(0,n,0), &\text{if } E_{\rm{max}} = E_1, \\
	(0,0,n), &\text{if } E_{\rm{max}} = E_2,
	\end{cases}
\end{equation}
so the multiplicity in both cases is $3^n$.

Using that $F = -\beta^{-1} \log \Z$, the first elbow coordinates in the two cases are now given by
\begin{equation}\label{eq:first_elb_coords}
	(x_0, L_0) =
	\begin{cases}
		\left(\frac{1}{3^n} e^{-n\beta (E_s - F)}, v^n \right), &\epsilon \leq \epsilon_\star \vspace{10pt}\\
		\left( e^{-n\beta (E_{\rm{max}}-F)}, (3u)^n \right). &\epsilon > \epsilon_\star
	\end{cases}
\end{equation}

\ddd{[Hmmm this needs an additional assumption on the output state to re-run the same analysis. Very annoying.]}\nick{What additional assumption beyond that the output state can be written as a Gibbs state for some $H'$?}
Similarly, considering the output magic state with respect to state $\sigma'$, the image of equilibrium state $\sigma$ under the magic protocol, we get output Lorenz curve coordinates,
\begin{equation}\label{eq:transformed_first_elb_coords}
	(x'_0, L'_0) =
	\begin{cases}
		\left(\frac{1}{3^{n'}} e^{-n\beta (E'_s - F')}, v(\epsilon')^{n'} \right), &\epsilon' \leq \epsilon'_\star \vspace{10pt}\\
		\left( e^{-n'\beta (E'_{\rm{max}}\hspace{-2.5pt}-F')}, (3u(\epsilon'))^{n'} \right), &\epsilon' > \epsilon'_\star
	\end{cases}
\end{equation}
There are four combinations of coordinates, depending on the noise parameters $\epsilon, \epsilon'$ for the input and output states.
In each of these combinations, we simply use the first elbow constraint, as described in~\cref{app:elb_constraints}, which leads to the bounds in the statement of the theorem.
\end{proof}

>>>>

The Wigner distribution of $n$ copies of the state is obtained by the $n$-fold product of the single copy distribution, and the Lorenz curve $L_{\rho|\sigma}(x)$ for $\rho = \rho_S(\epsilon)^{\otimes n}$ and $\sigma = (\I/3)^{\otimes n}$ is readily computed for any $n$ and $\epsilon$, and the full details are provided in~\cref{app:lcsu_technical}.
We highlight that the non-trivial element that arises for quasi-distributions is that the product of negative components can generate a positive $n$-fold component, so the ordering of the components of \emph{rescaled Wigner distribution} $W_{\rho |\sigma}(\bmx) \coloneqq W_{\rho}(\bmx)/W_{\sigma}(\bmx)$ depends on whether $-v(\epsilon)$ appears to an even or odd power in the expression of~\cref{eq:wigu}.

@@

The bounds also depend on the overlaps between the Hamiltonian eigenbases and the computational basis $\{\ket{k}\}$, where $k \in \mathbb{Z}_d$.
For a reference state $\tau$ with associated Hamiltonian $H$ denote the energy eigenbasis by $\{\ket{\phi_k}\}$ and define the \emph{reference partition function $\zeta$} as
\begin{equation}
	\zeta \coloneqq \sum_{k \in \mathbb{Z}_d} \alpha_k e^{-\beta E_k},
\end{equation}
where the overlap coefficients are
\begin{equation}
	\alpha_k \coloneqq \sum_{r \in \mathbb{Z}_d} \braket{\phi_k}{-r}\braket{r}{\phi_k}.
\end{equation}
Coefficients $\alpha_k$ can be negative when the Hamiltonian $H$ does not admit a Wigner-positive decomposition. 
However, it is simple to see that $\zeta$ is always positive. \nick{because it is proportional to $W_{\tau}(0,0) > 0$ for full-rank, stabilizer $\tau$ - this argument is a bit cyclic}
We can then naturally associate to the reference state $\tau$ a new form of free energy which is specific to the protocol, which we call the \emph{reference free energy $\Phi$} and define via
\begin{equation}
	\Phi \coloneqq -\beta^{-1} \log \zeta.
\end{equation}
For the output reference state $\sigma$ with associated Hamiltonian $H'$, we can denote the energy eigenbasis by $\{\ket{\phi'_k}\}$ and similarly define the reference partition function $\zeta'$ and reference free energy $\Phi'$.

@@

Then, the distillation rate $R = m/n$ of the magic protocol is bounded as:
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\Phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\Phi' - F')},
\end{equation}
where $F, \Phi$ are the free energy and reference free energy defined for state $\tau$, while $F', \Phi'$ are defined for state $\sigma$.
\end{theorem}
\begin{proof}	
For the sake of clarity, we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_m \coloneqq \rho_S(\epsilon')^{\otimes m}$, $\tau_n \coloneqq \tau^{\otimes m}$ and $\sigma_m \coloneqq \sigma^{\otimes m}$.

To establish the distillation bound we consider the distillation protocol that gives $\E( \rho_n) = \rho_m$ for the magic states. 
We then consider that the protocol transforms the reference equilibrium state as $\E( \tau_n) = \sigma_m$.
Since $\tau$ and $\sigma$ are assumed to be full rank stabilizer states, they have a strictly positive Wigner distribution, while, in contrast, the input and output magic states generally have quasi-probability Wigner distributions. 
For any such protocol we therefore have that
\begin{equation}
	( W_{\rho_n}(\bmx), W_{\tau_n}(\bmx) ) \succ ( W_{\rho'_m}(\bmx), W_{\sigma_m}(\bmx) ),
\end{equation}
or, equivalently, in terms of the relevant Lorenz curves,
\begin{equation}
	L_{\rho_n |\tau_n}(x) \ge L_{\rho'_m |\sigma_m}(x) \mbox{ for all } x.
\end{equation}

We define the rescaled Wigner distribution $W_{\rho | \tau}(\x) \coloneqq W_\rho(\x)/W_\tau(\x)$, which is always well-defined since $\tau$ is full-rank. 
Due to the multiplicative property of the Wigner distribution, the rescaled distribution is also multiplicative in the sense that
\begin{equation}
	W_{\rho \otimes \rho' | \tau \otimes \tau'} (\x_1 \oplus \x_2) = W_{\rho | \tau}(\x_1)W_{\rho' | \tau'}(\x_2),
\end{equation}
for any states $\rho, \rho'$ and any full-rank stabilizer states $\tau, \tau'$.
Therefore we have that
\begin{align}
	W_{\rho_n |\tau_n} (\x) &= \prod_{i=1}^n W_{\rho|\tau}(\x_i)\\
	W_{\rho_n } (\x) &= \prod_{i=1}^n W_\rho(\x_i)
\end{align}
where $\x = \oplus_{i=1}^n \x_i \in \mathbb{Z}_3^n$ is the phase space point for the full system in terms of those of the individual subsystems.

The points defining the Lorenz curve $L_{\rho_n |\tau_n}(x)$ are obtained from sorting the components of $W_{\rho_n |\tau_n}(\x)$ in non-increasing order and then computing the partial sums of $W_{\rho_n |\tau_n}(\pi(\x))$ where $\pi$ is the permutation that realises the sorting. 
However, similarly to the unital fragment analysis, we aim to use the constraint that is obtained by considering the line segment connecting the origin to the first elbow of both Lorenz curves.
The proof of this constraint is provided in~\cref{app:elb_constraints}.
Therefore, we focus on finding the coordinates of some point on this line segment of the Lorenz curves.

We find the coordinates for the input Lorenz curve, but the exact same analysis can be applied on the output Lorenz curve.
The Wigner distribution of a single noisy Strange state consists of 1 negative component $W_{\rho}(0,0) = -v(\epsilon)$ and 8 positive components $W_{\rho}(\bmx) = u(\epsilon)$ for $\bmx \neq \bmo$.
The Wigner distribution of the full-rank, stabilizer equilibrium state $\tau$ is $W_{\tau}(\bmx) > 0$ for all $\bmx \in \mathbb{Z}_3^2$.
We keep the components completely general.

Assume that the smallest component of the equilibrium state distribution is $W_{\tau}(i,j)$.
We then make use of the freedom to apply Clifford unitaries in order to perform the  Clifford operation
\begin{equation}
	\tau \longrightarrow X^{-i}Z^{-j}\ \tau\ Z^{j}X^{i},
\end{equation}
which permutes the smallest component to the origin of the phase space, $W_{\tau}(i,j) \longrightarrow W_{\tau}(0,0)$.
This operation would inevitably affect the magic distillation protocol, but since we consider the state distillation modulo Clifford operations, we are allowed to perform the reverse Clifford operation keeping the magic distillation process as is. \nick{I don't understand this too well, how are we allowed to permute $W_\tau$ while keeping $W_\rho$ intact?}

The rescaled distribution components are in general given by
\begin{equation}
	W_{\rho_n|\tau_n} = \left(\frac{-v}{W_{\tau}(0,0)}\right)^{i_{\bmo}} \prod_{\bmx \neq \bmo} \left(\frac{u}{W_{\tau}(\bmx)}\right)^{i_{\bmx}},
\end{equation}
where the (integer) indices obey the following conditions:
\begin{align}
&0 \leq i_{\bmx} \leq n \mbox{ for all } \bmx \in \mathbb{Z}_3^2, \nonumber \\
&\sum_{\bmx \in \mathbb{Z}_3^2} i_{\bmx} = n.
\end{align}
It is now easy to find the largest rescaled component.
Firstly, note that $n$ is even, so we require that $i_{\bmo} \in \{0,2,\dots,n\}$ for the component to be positive.
Then, we have that $v \geq u$ because $\epsilon \leq 3/7$, and we have already ensured that $W_{\tau}(0,0) \leq W_{\tau}(\bmx)$ for all $\bmx \in \mathbb{Z}_3^2$.
Therefore, the largest rescaled component occurs when $i_{\bmo} = n$ and $i_{\bmx} = 0$ for $\bmx \neq \bmo$ and is equal to $(v/W_{\tau}(0,0))^n$.
Accordingly, the coordinates of the first Lorenz curve point after the origin are given by
\begin{equation}
	(x_0, L_0) = ((W_{\tau}(0,0))^n, v^n)
\end{equation}

Given a Hamiltonian decomposition,
\begin{equation}
	H = \sum_{k \in \mathbb{Z}_3} E_k\ketbra{\phi_k}{\phi_k},
\end{equation}
we can re-express the coordinate location, by writing
\begin{align}
	W_{\tau}(0,0) &= \frac{1}{3\Z}\tr\left[ A_{0,0}e^{\beta H} \right] \nonumber\\
	&= \frac{e^{\beta F}}{3} \tr\left[ \sum_{r \in \mathbb{Z}_3} \ketbra{-r}{r}\sum_{k \in \mathbb{Z}_3} e^{\beta E_k}\ketbra{\phi_k}{\phi_k} \right] \\
	&= \frac{e^{\beta F}}{3} \sum_{r \in \mathbb{Z}_3} \alpha_k e^{\beta E_k}
	= \frac{e^{\beta F}}{3} \zeta \\
	&= \frac{e^{\beta (F - \Phi)}}{3},
\end{align}
where $\alpha_k, \zeta$ and $\Phi$ are as defined prior to the statement of the theorem.
Finally, the coordinates can be expressed as
\begin{equation}
	(x_0, L_0) = \left( \frac{e^{n\beta (F - \Phi)}}{3^n}, v(\epsilon)^n \right).
\end{equation}
In the same manner, we can derive the first point coordinates of the output Lorenz curve as
\begin{equation}
	(x'_0, L'_0) = \left( \frac{e^{n\beta (F' - \Phi')}}{3^n}, v(\epsilon')^n \right).
\end{equation}

If the largest rescaled component of a state is distinct with no multiplicities, then these coordinates correspond to the first elbow of the corresponding Lorenz curve, whereas if it appears multiple times, then the coordinates derived correspond to a point on the interior of the line segment connecting the origin to the first elbow.
In both cases, the bound derived remains the same, as is clear by its derivation in~\cref{app:elb_constraints}, and it is given by $L_0 / x_0 \geq L'_0 / x'_0$, which can be directly rearranged into the stated bound of the theorem.

\end{proof}

The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

\nick{FIX} As a special case of the result, suppose that $H=H'$, implying that the protocol leaves the equilibrium state unchanged, hence it corresponds to a Gibbs-preserving map~\cite{faist_2015}. In the more general case, $\Delta F \coloneqq F' - F \ne 0$ and we have a free energy difference between the initial and final states, with respect to the relevant Hamiltonians. This is similar to the free energy differences that occur in fluctuation relations, where the final state of the system can be arbitrarily far from equilibrium, yet a free energy term for the final system still arises.

\nick{FIX} The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.45]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
